---
title: "AI-Augmented Actuarial Reporting Pipeline"
author: "Rohith Narayanan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    code_folding: hide
runtime: shiny
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```
```{r load-libraries, include=FALSE}
suppressPackageStartupMessages({
  library(ChainLadder)
  library(dplyr)
  library(tidyr)
  library(knitr)
  library(kableExtra)
  library(ggplot2)
  library(shiny)
  library(httr)
  library(jsonlite)
})
```

# Executive Summary

This interactive report demonstrates an AI-augmented actuarial claims reserving analysis using the Chain Ladder method with scenario analysis, threshold-based uncertainty classification, and detailed pattern recognition. The analysis includes method suitability assessment and comparison with alternative approaches where appropriate.

# Data Input Section
```{r data-input, echo=FALSE}
fluidRow(
  column(12,
    wellPanel(
      h3("Data Input Options"),
      radioButtons("data_source", "Select Data Source:",
                   choices = c("Use Default GenIns Dataset" = "default",
                              "Upload CSV File" = "upload"),
                   selected = "default"),
      
      conditionalPanel(
        condition = "input.data_source == 'upload'",
        fileInput("file_upload", "Choose CSV File",
                  accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv")),
        helpText("CSV should have origin years as rows and development periods as columns.")
      ),
      
      hr(),
      h4("Scenario Analysis Configuration"),
      sliderInput("severity_shock", "Severity Shock (%):", 
                  min = -20, max = 20, value = 10, step = 5),
      sliderInput("dev_speed_shock", "Development Speed Shock (%):", 
                  min = -10, max = 10, value = 5, step = 2.5),
      checkboxInput("run_scenarios", "Run Scenario Analysis", value = TRUE),
      
      hr(),
      h4("AI Configuration"),
      selectInput("ai_provider", "Choose AI Provider:",
                  choices = c("Ollama (FREE - Local)" = "ollama",
                            "Anthropic Claude (Paid API)" = "anthropic",
                            "No AI (Skip Analysis)" = "none"),
                  selected = "ollama"),
      
      conditionalPanel(
        condition = "input.ai_provider == 'ollama'",
        div(
          style = "background-color: #d4edda; padding: 15px; border-radius: 5px; margin: 10px 0;",
          h5(style = "color: #155724; margin-top: 0;", "✓ FREE Local AI with Ollama"),
          p(style = "color: #155724; margin-bottom: 5px;", strong("Quick Setup:")),
          tags$ol(
            style = "color: #155724; margin-bottom: 10px;",
            tags$li("Download: ", tags$a(href = "https://ollama.com/download", target = "_blank", "ollama.com/download")),
            tags$li("Open Terminal/PowerShell and run:", tags$br(), 
                   tags$code("ollama serve"), tags$br(),
                   tags$small("(Keep this terminal window open!)")),
            tags$li("Open a NEW terminal and run:", tags$br(),
                   tags$code("ollama pull llama3.2")),
            tags$li("Click 'Test Ollama Connection' below before running analysis")
          )
        ),
        actionButton("test_ollama", "Test Ollama Connection", class = "btn-info"),
        textOutput("ollama_status"),
        br(),
        textInput("ollama_url", "Ollama Server URL:", 
                  value = "http://localhost:11434",
                  placeholder = "http://localhost:11434"),
        selectInput("ollama_model", "Select Model:",
                    choices = c("llama3.2" = "llama3.2",
                              "llama3.1" = "llama3.1", 
                              "llama2" = "llama2",
                              "mistral" = "mistral",
                              "phi3" = "phi3"),
                    selected = "llama3.2")
      ),
      
      conditionalPanel(
        condition = "input.ai_provider == 'anthropic'",
        div(
          style = "background-color: #fff3cd; padding: 15px; border-radius: 5px; margin: 10px 0;",
          p(style = "color: #856404; margin-bottom: 5px;", "⚠️ Requires paid API credits"),
          p(style = "color: #856404; margin-bottom: 0;", "Get your API key from: ", 
            tags$a(href = "https://console.anthropic.com/settings/keys", target = "_blank", 
                   "console.anthropic.com"))
        ),
        passwordInput("api_key_input", "Enter Anthropic API Key:", 
                      placeholder = "sk-ant-...")
      ),
      
      conditionalPanel(
        condition = "input.ai_provider == 'none'",
        div(
          style = "background-color: #e7f3ff; padding: 15px; border-radius: 5px; margin: 10px 0;",
          p(style = "color: #004085;", "ℹ️ Analysis will run without AI-generated insights.")
        )
      ),
      
      hr(),
      actionButton("run_analysis", "Run Analysis", class = "btn-primary btn-lg"),
      hr(),
      textOutput("data_status")
    )
  )
)
```
```{r test-ollama-connection}
observeEvent(input$test_ollama, {
  
  withProgress(message = 'Testing Ollama connection...', value = 0, {
    
    tryCatch({
      response <- GET(
        url = paste0(input$ollama_url, "/api/tags"),
        timeout(10)
      )
      
      if(status_code(response) == 200) {
        content_data <- content(response, "parsed")
        models <- sapply(content_data$models, function(x) x$name)
        
        if(input$ollama_model %in% models || any(grepl(input$ollama_model, models))) {
          output$ollama_status <- renderText({
            paste("✅ SUCCESS! Ollama is running and model", input$ollama_model, "is available.")
          })
          showNotification("Ollama connection successful!", type = "message", duration = 5)
        } else {
          output$ollama_status <- renderText({
            paste0("⚠️ Ollama is running but model '", input$ollama_model, 
                   "' not found. Available models: ", paste(models, collapse = ", "),
                   "\n\nRun in terminal: ollama pull ", input$ollama_model)
          })
          showNotification(paste("Model", input$ollama_model, "not found. Please pull it first."), 
                         type = "warning", duration = 10)
        }
      } else {
        output$ollama_status <- renderText({
          "❌ Ollama server responded but with an error. Make sure 'ollama serve' is running."
        })
      }
      
    }, error = function(e) {
      output$ollama_status <- renderText({
        paste0("❌ Cannot connect to Ollama at ", input$ollama_url, 
               "\n\nMake sure you run 'ollama serve' in terminal first!\n\n",
               "Error: ", e$message)
      })
      showNotification("Ollama connection failed! Run 'ollama serve' in terminal.", 
                      type = "error", duration = 10)
    })
  })
})
```
```{r reactive-data}
triangle_data <- reactiveVal(NULL)
analysis_triggered <- reactiveVal(FALSE)
user_api_key <- reactiveVal("")

observeEvent(input$api_key_input, {
  user_api_key(input$api_key_input)
})

observeEvent(input$run_analysis, {
  
  if(input$data_source == "default") {
    data(GenIns)
    triangle_data(GenIns)
    analysis_triggered(TRUE)
    
  } else if(input$data_source == "upload") {
    req(input$file_upload)
    
    tryCatch({
      uploaded <- read.csv(input$file_upload$datapath, row.names = 1)
      triangle_data(as.triangle(as.matrix(uploaded)))
      analysis_triggered(TRUE)
    }, error = function(e) {
      showNotification(paste("Error loading file:", e$message), type = "error")
      analysis_triggered(FALSE)
    })
  }
})

output$data_status <- renderText({
  if(!analysis_triggered()) {
    "No data loaded. Please select a data source and click 'Run Analysis'."
  } else {
    paste("✓ Data loaded successfully! Triangle dimensions:", 
          nrow(triangle_data()), "x", ncol(triangle_data()))
  }
})
```
```{r conditional-sections}
output$show_sections <- reactive({
  analysis_triggered()
})
outputOptions(output, "show_sections", suspendWhenHidden = FALSE)
```

# Method Selection & Suitability Assessment
```{r method-selection-section}
conditionalPanel(
  condition = "output.show_sections",
  uiOutput("method_selection_display")
)
```
```{r assess-method-suitability}
assess_cl_suitability <- function(triangle, cl_model, cl_summary) {
  
  # Calculate maturity for each origin year
  paid_by_year <- cl_model$FullTriangle[, ncol(cl_model$FullTriangle)]
  ultimate_by_year <- cl_summary$ByOrigin$Ultimate
  pct_developed <- (paid_by_year / ultimate_by_year) * 100
  
  immature_years <- sum(pct_developed < 30, na.rm = TRUE)
  
  # Check CV volatility
  cv_by_year <- (cl_summary$ByOrigin$Mack.S.E / cl_summary$ByOrigin$IBNR) * 100
  high_cv_years <- sum(cv_by_year > 30, na.rm = TRUE)
  
  # Check development factor stability
  dev_factors <- as.vector(cl_model$f)
  df_cv <- sd(dev_factors, na.rm = TRUE) / mean(dev_factors, na.rm = TRUE) * 100
  
  # Red flags
  red_flags <- list()
  
  if(immature_years > 2) {
    red_flags <- c(red_flags, list(list(
      flag = "Immature Data",
      severity = "Medium",
      description = paste0(immature_years, " origin year(s) are less than 30% developed"),
      recommendation = "Consider Bornhuetter-Ferguson for recent accident years"
    )))
  }
  
  if(high_cv_years > 1) {
    red_flags <- c(red_flags, list(list(
      flag = "High Uncertainty",
      severity = "Medium",
      description = paste0(high_cv_years, " origin year(s) have CV > 30%"),
      recommendation = "Review underlying data quality; consider segmentation"
    )))
  }
  
  if(df_cv > 25) {
    red_flags <- c(red_flags, list(list(
      flag = "Volatile Development Factors",
      severity = "High",
      description = paste0("Development factor CV = ", round(df_cv, 1), "% indicates instability"),
      recommendation = "Investigate process changes or consider alternative methods"
    )))
  }
  
  # Check for exposure changes (simplified - based on paid trends)
  if(nrow(triangle) >= 3) {
    recent_paid <- sum(triangle[nrow(triangle), ], na.rm = TRUE)
    earlier_paid <- sum(triangle[max(1, nrow(triangle)-2), ], na.rm = TRUE)
    
    if(abs(recent_paid / earlier_paid - 1) > 0.5) {
      red_flags <- c(red_flags, list(list(
        flag = "Potential Exposure Change",
        severity = "High",
        description = "Significant change in paid claims between recent origin years",
        recommendation = "Verify if portfolio exposure has materially changed; if yes, use BF method"
      )))
    }
  }
  
  list(
    red_flags = red_flags,
    immature_years = immature_years,
    high_cv_years = high_cv_years,
    df_cv = df_cv,
    pct_developed = pct_developed
  )
}
```
```{r display-method-selection}
output$method_selection_display <- renderUI({
  req(analysis_triggered())
  req(triangle_data())
  
  triangle <- triangle_data()
  cl_model <- MackChainLadder(triangle, est.sigma = "Mack")
  cl_summary <- summary(cl_model)
  
  suitability <- assess_cl_suitability(triangle, cl_model, cl_summary)
  
  div(
    h2("Method Selection & Suitability Assessment"),
    
    div(
      style = "background-color: #e7f3ff; padding: 20px; border-left: 4px solid #3498db; border-radius: 5px; margin-bottom: 20px;",
      h4(style = "margin-top: 0; color: #004085;", "Chain Ladder Method Assumptions"),
      p("The Chain Ladder method is appropriate when the following conditions hold:"),
      tags$ul(
        tags$li(strong("Stable development patterns: "), "Past claim development is representative of future patterns"),
        tags$li(strong("Homogeneous portfolio: "), "No material changes in business mix, policy terms, or coverage"),
        tags$li(strong("Consistent claims handling: "), "Settlement practices and claims processes remain stable"),
        tags$li(strong("Proportional exposure: "), "No significant changes in premium volume or portfolio size"),
        tags$li(strong("Sufficient maturity: "), "Adequate development history with low volatility")
      ),
      
      p(style = "margin-top: 15px; margin-bottom: 0;",
        strong("Important Note: "), 
        "If exposure volumes, claim handling practices, policy conditions, or business mix have materially changed, ",
        "Chain Ladder results may be biased. Alternative methods such as Bornhuetter-Ferguson should be considered in such cases.")
    ),
    
    if(length(suitability$red_flags) > 0) {
      div(
        style = "background-color: #fff3cd; padding: 20px; border-left: 4px solid #ffc107; border-radius: 5px; margin-bottom: 20px;",
        h4(style = "margin-top: 0; color: #856404;", 
           "⚠️ Method Suitability Warnings (", length(suitability$red_flags), " issue", 
           ifelse(length(suitability$red_flags) > 1, "s", ""), " detected)"),
        
        tags$table(
          class = "table table-striped",
          style = "margin-top: 15px;",
          tags$thead(
            tags$tr(
              tags$th("Issue", style = "width: 25%;"),
              tags$th("Severity", style = "width: 15%;"),
              tags$th("Description", style = "width: 35%;"),
              tags$th("Recommendation", style = "width: 25%;")
            )
          ),
          tags$tbody(
            lapply(suitability$red_flags, function(flag) {
              severity_color <- ifelse(flag$severity == "High", "#dc3545", 
                                      ifelse(flag$severity == "Medium", "#ffc107", "#28a745"))
              tags$tr(
                tags$td(flag$flag),
                tags$td(tags$span(flag$severity, style = paste0("color: ", severity_color, "; font-weight: bold;"))),
                tags$td(flag$description),
                tags$td(flag$recommendation)
              )
            })
          )
        )
      )
    } else {
      div(
        style = "background-color: #d4edda; padding: 20px; border-left: 4px solid #28a745; border-radius: 5px; margin-bottom: 20px;",
        h4(style = "margin-top: 0; color: #155724;", "✓ Chain Ladder Appears Suitable"),
        p("No major red flags detected. The data appears suitable for Chain Ladder analysis based on automated checks.")
      )
    },
    
    h4("Alternative Methods for Comparison"),
    div(
      style = "background-color: #f8f9fa; padding: 20px; border-radius: 5px;",
      tags$table(
        class = "table table-bordered",
        tags$thead(
          tags$tr(
            tags$th("Problem", style = "width: 30%;"),
            tags$th("Better Method", style = "width: 25%;"),
            tags$th("Why It Helps", style = "width: 45%;")
          )
        ),
        tags$tbody(
          tags$tr(
            tags$td("Exposure change"),
            tags$td(strong("Bornhuetter-Ferguson (BF)")),
            tags$td("Separates a priori loss expectations from observed development; reduces bias from volume changes")
          ),
          tags$tr(
            tags$td("Immature data"),
            tags$td(strong("BF / Cape Cod")),
            tags$td("Reduces reliance on unstable early development factors; blends prior knowledge with data")
          ),
          tags$tr(
            tags$td("Claims process change"),
            tags$td(strong("BF or segmentation")),
            tags$td("Less sensitive to recent process distortions; segments isolate structural breaks")
          ),
          tags$tr(
            tags$td("Volatile large losses"),
            tags$td(strong("Large loss carve-out")),
            tags$td("Prevents individual claims from distorting development patterns")
          ),
          tags$tr(
            tags$td("Changing mix of business"),
            tags$td(strong("Segmented triangles")),
            tags$td("Restores homogeneity by analyzing similar cohorts separately")
          ),
          tags$tr(
            tags$td("High uncertainty"),
            tags$td(strong("Mack + stress testing")),
            tags$td("Quantifies and communicates reserve risk explicitly")
          )
        )
      ),
      
      p(style = "margin-top: 15px; font-style: italic; color: #666;",
        strong("Note: "), "This analysis uses the Mack Chain Ladder as the primary method. ",
        "In practice, actuaries should compare multiple methods (e.g., Chain Ladder vs. BF) and select ",
        "the most appropriate based on data characteristics, business context, and professional judgment. ",
        "The methods above provide a framework for such comparisons.")
    )
  )
})
```

# Assumptions & Methodology
```{r assumptions-section}
conditionalPanel(
  condition = "output.show_sections",
  uiOutput("assumptions_display")
)
```
```{r display-assumptions}
output$assumptions_display <- renderUI({
  req(analysis_triggered())
  
  HTML(paste0(
    "<div style='background-color: #f8f9fa; padding: 20px; border-left: 4px solid #5bc0de; border-radius: 5px;'>",
    "<h3 style='margin-top: 0; color: #31708f;'>Methodology & Key Assumptions</h3>",
    
    "<h4>Chain Ladder Method</h4>",
    "<ul>",
    "<li><strong>Development Factor Calculation:</strong> Volume-weighted age-to-age factors</li>",
    "<li><strong>Standard Error Estimation:</strong> Mack (1993) method for measuring reserve uncertainty</li>",
    "<li><strong>Independence Assumption:</strong> Accident years develop independently</li>",
    "<li><strong>Projection Method:</strong> Full triangle projection to ultimate using iterative application of development factors</li>",
    "<li><strong>Critical Limitation:</strong> Assumes past development patterns remain representative of future development</li>",
    "</ul>",
    
    "<h4>Data Specifications</h4>",
    "<ul>",
    "<li><strong>Data Cut-off Date:</strong> ", format(Sys.Date(), '%B %d, %Y'), "</li>",
    "<li><strong>Currency:</strong> Values in thousands (unless otherwise specified)</li>",
    "<li><strong>Triangle Type:</strong> Paid claims development triangle</li>",
    "<li><strong>No Manual Adjustments:</strong> Base case uses raw data without manual factor smoothing</li>",
    "</ul>",
    
    "<h4>Uncertainty Classification Thresholds</h4>",
    "<table class='table table-bordered' style='width: 70%; margin-top: 10px;'>",
    "<thead><tr><th>Coefficient of Variation (CV)</th><th>Uncertainty Level</th><th>Interpretation</th></tr></thead>",
    "<tbody>",
    "<tr><td>< 10%</td><td><span style='color: #28a745; font-weight: bold;'>Low</span></td><td>High confidence in estimate</td></tr>",
    "<tr><td>10% - 20%</td><td><span style='color: #ffc107; font-weight: bold;'>Moderate</span></td><td>Reasonable confidence, some variability expected</td></tr>",
    "<tr><td>> 20%</td><td><span style='color: #dc3545; font-weight: bold;'>High</span></td><td>Significant uncertainty, requires detailed review</td></tr>",
    "</tbody></table>",
    
    "<h4>Development Factor Interpretation</h4>",
    "<table class='table table-bordered' style='width: 80%; margin-top: 10px;'>",
    "<thead><tr><th>Development Factor</th><th>Classification</th><th>Interpretation</th></tr></thead>",
    "<tbody>",
    "<tr><td>> 1.20</td><td><span style='color: #dc3545; font-weight: bold;'>High</span></td><td>May indicate immature claims, process changes, or reporting delays — not necessarily higher ultimate losses</td></tr>",
    "<tr><td>1.05 - 1.20</td><td><span style='color: #ffc107; font-weight: bold;'>Moderate</span></td><td>Normal development pattern for mid-stage claims</td></tr>",
    "<tr><td>< 1.05</td><td><span style='color: #28a745; font-weight: bold;'>Low/Mature</span></td><td>Claims nearing full development; minimal future movement expected</td></tr>",
    "</tbody></table>",
    
    "<p style='background-color: #fff3cd; padding: 12px; border-radius: 5px; margin-top: 15px;'>",
    "<strong>⚠️ Important Actuarial Note:</strong> ",
    "Development factors reflect <em>cumulative development speed</em>, not claim magnitude. ",
    "A declining development factor pattern may indicate faster claim settlement, changes in reporting practices, ",
    "improved claims handling efficiency, or reinsurance effects — rather than purely lower claim frequency or severity. ",
    "Conversely, increasing factors may signal reporting delays or process slowdowns, not necessarily adverse development.",
    "</p>",
    
    "<h4>Scenario Analysis Design</h4>",
    "<ul>",
    "<li><strong>Approach:</strong> Illustrative, deterministic stress testing</li>",
    "<li><strong>Severity Shock:</strong> Uniform percentage change applied to all claim amounts across all origin years</li>",
    "<li><strong>Development Speed Shock:</strong> Uniform percentage change applied to all development factors</li>",
    "<li><strong>Limitation:</strong> Scenarios are symmetric and do not reflect year-specific or asymmetric shocks</li>",
    "<li><strong>Enhancement Opportunity:</strong> More advanced stress testing could incorporate stochastic simulations, ",
    "asymmetric shocks, or origin-year-specific adjustments based on exposure or claims experience</li>",
    "</ul>",
    
    "<h4>Known Limitations & Caveats</h4>",
    "<div style='background-color: #ffe5e5; padding: 15px; border-radius: 5px; margin-top: 10px;'>",
    "<p style='margin-bottom: 8px;'><strong>This analysis does NOT account for:</strong></p>",
    "<ul style='margin-bottom: 0;'>",
    "<li>Material changes in exposure volume (premiums doubled/halved, portfolio growth/shrinkage)</li>",
    "<li>Changes in claims handling systems or settlement speed (new systems, process reforms)</li>",
    "<li>Changes in business mix (new products, different policy terms, geographic expansion)</li>",
    "<li>Large loss concentration or catastrophe events that may distort development patterns</li>",
    "<li>Regulatory changes, legal reforms, or external economic shocks</li>",
    "</ul>",
    "<p style='margin-top: 10px; margin-bottom: 0; font-style: italic;'>",
    "If any of these conditions are present, the Chain Ladder estimates may be materially biased. ",
    "Review the Method Suitability Assessment section and consider alternative approaches.",
    "</p>",
    "</div>",
    
    "<h4>Reproducibility & Audit Trail</h4>",
    "<p style='margin-top: 15px; font-style: italic; color: #666;'>",
    "All calculations follow the Mack Chain Ladder methodology as implemented in the ChainLadder R package (version ", 
    as.character(packageVersion("ChainLadder")), "). ",
    "Results are fully reproducible given the same input data. Development factors, standard errors, and ultimate projections ",
    "are calculated according to Mack (1993) formulas. All code is available for audit and can be provided upon request.",
    "</p>",
    "</div>"
  ))
})
```

# Input Data
```{r show-input-data, eval=TRUE}
conditionalPanel(
  condition = "output.show_sections",
  
  div(
    h2("Input Data"),
    
    h3("Claims Triangle"),
    uiOutput("triangle_display"),
    
    h3("Data Summary"),
    uiOutput("summary_display")
  )
)
```
```{r display-triangle}
output$triangle_display <- renderUI({
  req(analysis_triggered())
  req(triangle_data())
  
  triangle <- triangle_data()
  triangle_df <- as.data.frame(as.matrix(triangle))
  triangle_df <- cbind(Origin = rownames(triangle_df), triangle_df)
  rownames(triangle_df) <- NULL
  
  kable(triangle_df, 
        caption = "Claims Triangle (in thousands)",
        format.args = list(big.mark = ",", scientific = FALSE),
        align = c("l", rep("r", ncol(triangle)))) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE,
                  position = "center",
                  font_size = 12) %>%
    HTML()
})
```
```{r display-summary}
output$summary_display <- renderUI({
  req(analysis_triggered())
  req(triangle_data())
  
  triangle <- triangle_data()
  
  summary_stats <- data.frame(
    Metric = c("Origin Years", "Development Years", "Total Observations", 
               "Minimum Value", "Maximum Value", "Mean Value"),
    Value = c(
      nrow(triangle),
      ncol(triangle),
      sum(!is.na(triangle)),
      format(min(triangle, na.rm = TRUE), big.mark = ","),
      format(max(triangle, na.rm = TRUE), big.mark = ","),
      format(round(mean(as.matrix(triangle), na.rm = TRUE), 0), big.mark = ",")
    )
  )
  
  kable(summary_stats, 
        caption = "Triangle Summary Statistics",
        col.names = c("Metric", "Value"),
        align = c("l", "r")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE,
                  position = "center") %>%
    HTML()
})
```

# AI-Generated Insights
```{r ai-insights-section}
conditionalPanel(
  condition = "output.show_sections",
  uiOutput("ai_insights_display")
)
```
```{r pattern-analysis}
analyze_patterns <- function(triangle, cl_model, cl_summary) {
  
  total_ibnr <- cl_summary$Totals["IBNR", ]
  total_se <- cl_summary$Totals["Mack S.E", ]
  cv <- total_se / total_ibnr * 100
  latest <- cl_summary$Totals["Latest", ]
  ultimate <- cl_summary$Totals["Ultimate", ]
  reserve_ratio <- total_ibnr / latest * 100
  
  dev_factors <- as.vector(cl_model$f)
  high_dfs <- which(dev_factors > 1.20)
  moderate_dfs <- which(dev_factors >= 1.05 & dev_factors <= 1.20)
  low_dfs <- which(dev_factors < 1.05)
  
  ibnr_by_year <- cl_summary$ByOrigin$IBNR
  max_ibnr_idx <- which.max(ibnr_by_year[1:(length(ibnr_by_year)-1)])
  max_ibnr_year <- rownames(cl_summary$ByOrigin)[max_ibnr_idx]
  max_ibnr_value <- ibnr_by_year[max_ibnr_idx]
  max_ibnr_pct <- (max_ibnr_value / total_ibnr) * 100
  
  cv_by_year <- (cl_summary$ByOrigin$Mack.S.E / cl_summary$ByOrigin$IBNR) * 100
  high_uncertainty_years <- which(cv_by_year > 20)
  
  df_trend <- if(length(dev_factors) >= 3) {
    recent_avg <- mean(dev_factors[(length(dev_factors)-2):length(dev_factors)])
    early_avg <- mean(dev_factors[1:min(3, length(dev_factors))])
    if(recent_avg < early_avg * 0.9) "declining" else if(recent_avg > early_avg * 1.1) "increasing" else "stable"
  } else {
    "stable"
  }
  
  # Maturity assessment
  paid_by_year <- cl_model$FullTriangle[, ncol(cl_model$FullTriangle)]
  ultimate_by_year <- cl_summary$ByOrigin$Ultimate
  pct_developed <- (paid_by_year / ultimate_by_year) * 100
  immature_years <- sum(pct_developed < 30, na.rm = TRUE)
  
  list(
    total_ibnr = total_ibnr,
    total_se = total_se,
    cv = cv,
    latest = latest,
    ultimate = ultimate,
    reserve_ratio = reserve_ratio,
    dev_factors = dev_factors,
    high_dfs = high_dfs,
    moderate_dfs = moderate_dfs,
    low_dfs = low_dfs,
    max_ibnr_year = max_ibnr_year,
    max_ibnr_value = max_ibnr_value,
    max_ibnr_pct = max_ibnr_pct,
    high_uncertainty_years = high_uncertainty_years,
    df_trend = df_trend,
    immature_years = immature_years,
    pct_developed = pct_developed
  )
}
```
```{r ai-analysis-function}
get_ai_insights_ollama <- function(triangle, cl_model, cl_summary, patterns, suitability, ollama_url, model_name) {
  
  prompt <- paste0(
    "You are a senior actuary performing a detailed Chain Ladder analysis. Provide specific, quantitative insights with explicit pattern recognition.\n\n",
    
    "=== KEY METRICS ===\n",
    "Triangle: ", nrow(triangle), " origin years × ", ncol(triangle), " development periods\n",
    "Latest Paid Claims: $", format(round(patterns$latest, 0), big.mark = ","), "\n",
    "Ultimate Claims Projection: $", format(round(patterns$ultimate, 0), big.mark = ","), "\n",
    "IBNR Reserve: $", format(round(patterns$total_ibnr, 0), big.mark = ","), "\n",
    "Standard Error: $", format(round(patterns$total_se, 0), big.mark = ","), "\n",
    "Coefficient of Variation: ", round(patterns$cv, 2), "%\n",
    "Reserve-to-Paid Ratio: ", round(patterns$reserve_ratio, 1), "%\n\n",
    
    "=== UNCERTAINTY CLASSIFICATION ===\n",
    "Based on CV = ", round(patterns$cv, 2), "%, this falls into: ",
    ifelse(patterns$cv < 10, "LOW uncertainty (CV < 10%)", 
           ifelse(patterns$cv < 20, "MODERATE uncertainty (10% ≤ CV < 20%)", 
                  "HIGH uncertainty (CV ≥ 20%)")), "\n\n",
    
    "=== DEVELOPMENT FACTOR ANALYSIS ===\n",
    "Development Factors: ", paste(round(patterns$dev_factors, 4), collapse = ", "), "\n",
    "Trend: ", toupper(patterns$df_trend), "\n",
    "High DFs (>1.20) at periods: ", if(length(patterns$high_dfs) > 0) paste(patterns$high_dfs, collapse = ", ") else "None", "\n",
    "Moderate DFs (1.05-1.20) at periods: ", if(length(patterns$moderate_dfs) > 0) paste(patterns$moderate_dfs, collapse = ", ") else "None", "\n",
    "Low/Mature DFs (<1.05) at periods: ", if(length(patterns$low_dfs) > 0) paste(patterns$low_dfs, collapse = ", ") else "None", "\n",
    "IMPORTANT: Development factors reflect settlement speed, not claim size. ",
    if(patterns$df_trend == "declining") {
      "Declining DFs may indicate faster settlements, reporting improvements, or process efficiencies — not necessarily lower severities."
    } else if(patterns$df_trend == "increasing") {
      "Increasing DFs may signal reporting delays or slower settlements — not necessarily adverse claim development."
    } else {
      "Stable DFs suggest consistent claims handling patterns."
    }, "\n\n",
    
    "=== MATURITY ASSESSMENT ===\n",
    "Immature years (< 30% developed): ", patterns$immature_years, "\n",
    if(patterns$immature_years > 2) {
      "WARNING: Multiple immature years may make Chain Ladder unreliable. Consider Bornhuetter-Ferguson for recent years.\n"
    } else {
      "Maturity levels appear adequate for Chain Ladder.\n"
    }, "\n",
    
    "=== IBNR CONCENTRATION ===\n",
    "Largest IBNR: Origin Year ", patterns$max_ibnr_year, 
    " = $", format(round(patterns$max_ibnr_value, 0), big.mark = ","), 
    " (", round(patterns$max_ibnr_pct, 1), "% of total IBNR)\n",
    "High Uncertainty Years (CV > 20%): ", if(length(patterns$high_uncertainty_years) > 0) 
      paste(rownames(cl_summary$ByOrigin)[patterns$high_uncertainty_years], collapse = ", ") else "None", "\n\n",
    
    "=== METHOD SUITABILITY FLAGS ===\n",
    if(length(suitability$red_flags) > 0) {
      paste0("WARNINGS DETECTED (", length(suitability$red_flags), "):\n",
            paste(sapply(suitability$red_flags, function(f) 
              paste0("- ", f$flag, " (", f$severity, "): ", f$description)), collapse = "\n"), "\n\n")
    } else {
      "No method suitability warnings detected.\n\n"
    },
    
    "=== REQUIRED ANALYSIS ===\n",
    "Provide a professional actuarial analysis with:\n\n",
    
    "1. EXECUTIVE SUMMARY (3-4 sentences)\n",
    "   - State the total IBNR with context (absolute value + reserve ratio)\n",
    "   - Classify uncertainty level and explain why based on CV\n",
    "   - Mention any method suitability concerns if present\n",
    "   - Highlight the single most important finding\n\n",
    
    "2. DEVELOPMENT PATTERN INSIGHTS\n",
    "   - Explain the trend (", patterns$df_trend, ") with CORRECT actuarial interpretation\n",
    "   - Remember: DFs reflect development SPEED, not claim magnitude\n",
    "   - Flag any high DFs (>1.20) with period numbers and possible causes\n",
    "   - Note maturity indicators (low DFs at later periods)\n",
    "   - Discuss what the pattern suggests about claims handling or reporting\n\n",
    
    "3. RESERVE CONCENTRATION & DATA QUALITY\n",
    "   - Discuss origin year ", patterns$max_ibnr_year, " holding ", round(patterns$max_ibnr_pct, 1), "% of IBNR\n",
    "   - Flag any years with high uncertainty (CV > 20%)\n",
    "   - Assess concentration risk\n",
    "   - Comment on data maturity issues if present\n\n",
    
    "4. METHOD APPROPRIATENESS & ALTERNATIVES\n",
    "   - Is Chain Ladder appropriate for this data?\n",
    "   - Should Bornhuetter-Ferguson be considered? (especially if immature years present)\n",
    "   - Any concerns about exposure changes or mix shifts?\n\n",
    
    "5. ACTIONABLE RECOMMENDATIONS\n",
    "   - Specific next steps based on the patterns observed\n",
    "   - Data quality checks needed (if applicable)\n",
    "   - Monitoring priorities\n",
    "   - Alternative methods to consider\n\n",
    
    "Use specific numbers from the data above. Avoid generic statements. ",
    "Each point should reference actual values or thresholds. Keep total response under 450 words but be specific and quantitative. ",
    "Use professional actuarial language but explain technical points clearly."
  )
  
  tryCatch({
    test_response <- GET(url = paste0(ollama_url, "/api/tags"), timeout(5))
    
    if(status_code(test_response) != 200) {
      return("❌ Ollama server not responding. Run 'ollama serve' in terminal.")
    }
    
    response <- POST(
      url = paste0(ollama_url, "/api/generate"),
      body = toJSON(list(model = model_name, prompt = prompt, stream = FALSE), auto_unbox = TRUE),
      encode = "json",
      content_type_json(),
      timeout(120)
    )
    
    if(status_code(response) == 200) {
      content_data <- content(response, "parsed")
      return(content_data$response)
    } else {
      return(paste0("❌ Model '", model_name, "' not found. Run: ollama pull ", model_name))
    }
    
  }, error = function(e) {
    if(grepl("Timeout", e$message)) {
      return("⏱️ AI taking too long. Wait 1-2 min and retry, or select 'No AI'.")
    }
    return(paste0("❌ Error: ", e$message, "\n\nRun: ollama serve"))
  })
}

get_ai_insights_anthropic <- function(triangle, cl_model, cl_summary, patterns, suitability, api_key) {
  
  if(is.null(api_key) || api_key == "" || nchar(api_key) < 10) {
    return("⚠️ Please enter your Anthropic API key above.")
  }
  
  prompt <- paste0(
    "You are a senior actuary. Provide detailed, quantitative Chain Ladder analysis with correct actuarial interpretation.\n\n",
    
    "KEY METRICS:\n",
    "Triangle: ", nrow(triangle), " origin years × ", ncol(triangle), " periods\n",
    "Latest Paid: $", format(round(patterns$latest, 0), big.mark = ","), "\n",
    "Ultimate: $", format(round(patterns$ultimate, 0), big.mark = ","), "\n",
    "IBNR: $", format(round(patterns$total_ibnr, 0), big.mark = ","), "\n",
    "Std Error: $", format(round(patterns$total_se, 0), big.mark = ","), "\n",
    "CV: ", round(patterns$cv, 2), "% (", 
    ifelse(patterns$cv < 10, "LOW", ifelse(patterns$cv < 20, "MODERATE", "HIGH")), " uncertainty)\n",
    "Reserve/Paid: ", round(patterns$reserve_ratio, 1), "%\n\n",
    
    "DEVELOPMENT FACTORS: ", paste(round(patterns$dev_factors, 4), collapse = ", "), "\n",
    "Trend: ", toupper(patterns$df_trend), "\n",
    "CRITICAL: DFs reflect SETTLEMENT SPEED, not claim magnitude. ",
    if(patterns$df_trend == "declining") "Declining may = faster settlements/reporting, not lower claims." else 
      if(patterns$df_trend == "increasing") "Increasing may = delays, not adverse development." else "Stable = consistent processes.", "\n",
    "High DFs (>1.20): periods ", if(length(patterns$high_dfs) > 0) paste(patterns$high_dfs, collapse = ", ") else "None", "\n",
    "Immature years (<30% developed): ", patterns$immature_years, 
    if(patterns$immature_years > 2) " → Consider BF method" else "", "\n\n",
    
    "MAX IBNR: Year ", patterns$max_ibnr_year, " = $", format(round(patterns$max_ibnr_value, 0), big.mark = ","),
    " (", round(patterns$max_ibnr_pct, 1), "% of total)\n\n",
    
    if(length(suitability$red_flags) > 0) {
      paste0("METHOD WARNINGS (", length(suitability$red_flags), "):\n",
            paste(sapply(suitability$red_flags, function(f) paste0("- ", f$flag)), collapse = "\n"), "\n\n")
    } else {
      ""
    },
    
    "Provide:\n",
    "1. Executive summary with specific numbers and method concerns\n",
    "2. Development pattern insights (CORRECT interpretation of DFs)\n",
    "3. Reserve concentration risks and data maturity\n",
    "4. Method appropriateness (should we use BF instead?)\n",
    "5. Actionable recommendations\n",
    "Be specific and quantitative. Under 450 words."
  )
  
  tryCatch({
    response <- POST(
      url = "https://api.anthropic.com/v1/messages",
      add_headers(
        "x-api-key" = api_key,
        "anthropic-version" = "2023-06-01",
        "content-type" = "application/json"
      ),
      body = toJSON(list(
        model = "claude-sonnet-4-20250514",
        max_tokens = 1500,
        messages = list(list(role = "user", content = prompt))
      ), auto_unbox = TRUE),
      encode = "json"
    )
    
    if(status_code(response) == 200) {
      return(content(response, "parsed")$content[[1]]$text)
    } else {
      return(paste0("❌ API Error: ", status_code(response)))
    }
    
  }, error = function(e) {
    return(paste0("❌ Error: ", e$message))
  })
}
```
```{r display-ai-insights}
output$ai_insights_display <- renderUI({
  req(analysis_triggered())
  req(triangle_data())
  
  triangle <- triangle_data()
  cl_model <- MackChainLadder(triangle, est.sigma = "Mack")
  cl_summary <- summary(cl_model)
  
  patterns <- analyze_patterns(triangle, cl_model, cl_summary)
  suitability <- assess_cl_suitability(triangle, cl_model, cl_summary)
  
  if(input$ai_provider == "none") {
    return(
      wellPanel(
        h3("ℹ️ AI Analysis Skipped", style = "color: #666;"),
        p("You selected 'No AI'. The analysis has been completed without AI-generated insights.")
      )
    )
  }
  
  withProgress(message = 'Generating detailed AI insights...', value = 0.5, {
    
    if(input$ai_provider == "ollama") {
      insights <- get_ai_insights_ollama(triangle, cl_model, cl_summary, patterns, suitability,
                                         input$ollama_url, input$ollama_model)
      provider_label <- "🤖 AI-Powered Actuarial Analysis (FREE - Ollama Local)"
      provider_color <- "#27ae60"
    } else {
      insights <- get_ai_insights_anthropic(triangle, cl_model, cl_summary, patterns, suitability,
                                            user_api_key())
      provider_label <- "🤖 AI-Powered Actuarial Analysis (Claude API)"
      provider_color <- "#3498db"
    }
  })
  
  wellPanel(
    h3(provider_label, style = paste0("color: ", provider_color, ";")),
    HTML(paste0("<div style='background-color: #f8f9fa; padding: 20px; border-left: 4px solid ", 
                provider_color, "; border-radius: 5px;'>",
                "<pre style='white-space: pre-wrap; font-family: Arial, sans-serif; font-size: 14px; line-height: 1.6;'>",
                insights,
                "</pre></div>"))
  )
})
```

# Scenario Analysis
```{r scenario-section}
conditionalPanel(
  condition = "output.show_sections && input.run_scenarios",
  uiOutput("scenario_display")
)
```
```{r run-scenarios}
output$scenario_display <- renderUI({
  req(analysis_triggered())
  req(triangle_data())
  req(input$run_scenarios)
  
  triangle <- triangle_data()
  cl_base <- MackChainLadder(triangle, est.sigma = "Mack")
  
  triangle_sev <- triangle * (1 + input$severity_shock/100)
  cl_sev <- MackChainLadder(triangle_sev, est.sigma = "Mack")
  
  cl_speed <- cl_base
  cl_speed$f <- cl_base$f * (1 + input$dev_speed_shock/100)
  triangle_speed <- triangle
  for(i in 1:nrow(triangle)) {
    for(j in 1:(ncol(triangle)-1)) {
      if(!is.na(triangle[i,j]) && is.na(triangle[i,j+1])) {
        triangle_speed[i,j+1] <- triangle[i,j] * cl_speed$f[j]
      }
    }
  }
  cl_speed <- MackChainLadder(triangle_speed, est.sigma = "Mack")
  
  base_ibnr <- summary(cl_base)$Totals["IBNR", ]
  sev_ibnr <- summary(cl_sev)$Totals["IBNR", ]
  speed_ibnr <- summary(cl_speed)$Totals["IBNR", ]
  
  base_cv <- (summary(cl_base)$Totals["Mack S.E", ] / base_ibnr) * 100
  sev_cv <- (summary(cl_sev)$Totals["Mack S.E", ] / sev_ibnr) * 100
  speed_cv <- (summary(cl_speed)$Totals["Mack S.E", ] / speed_ibnr) * 100
  
  scenario_df <- data.frame(
    Scenario = c("Base Case", 
                 paste0("Severity +", input$severity_shock, "%"),
                 paste0("Dev Speed +", input$dev_speed_shock, "%")),
    IBNR = c(base_ibnr, sev_ibnr, speed_ibnr),
    Change_from_Base = c(0, sev_ibnr - base_ibnr, speed_ibnr - base_ibnr),
    Pct_Change = c(0, ((sev_ibnr/base_ibnr - 1) * 100), ((speed_ibnr/base_ibnr - 1) * 100)),
    CV = c(base_cv, sev_cv, speed_cv)
  )
  
  div(
    h2("Scenario Analysis Results"),
    
    div(
      style = "background-color: #fff3cd; padding: 15px; border-left: 4px solid #ffc107; border-radius: 5px; margin-bottom: 20px;",
      h5(style = "margin-top: 0; color: #856404;", "Scenario Design Limitations"),
      p("These scenarios are illustrative and apply uniform, symmetric, deterministic shocks across all origin years. ",
        "More advanced stress testing could incorporate asymmetric shocks, year-specific adjustments based on exposure changes, ",
        "or stochastic simulations to better reflect real-world uncertainty.")
    ),
    
    p(strong("Scenario Assumptions:")),
    tags$ul(
      tags$li("Severity Shock: ", input$severity_shock, "% uniform increase applied to all claim amounts across all origin years"),
      tags$li("Development Speed Shock: ", input$dev_speed_shock, "% uniform increase applied to all development factors")
    ),
    
    kable(scenario_df,
          caption = "IBNR Reserve Sensitivity Analysis",
          digits = c(0, 0, 0, 1, 2),
          format.args = list(big.mark = ","),
          col.names = c("Scenario", "IBNR Reserve", "Change ($)", "Change (%)", "CV (%)"),
          align = c("l", "r", "r", "r", "r")) %>%
      kable_styling(bootstrap_options = c("striped", "hover"),
                    full_width = FALSE,
                    position = "center") %>%
      row_spec(1, bold = TRUE, background = "#e7f3ff") %>%
      HTML(),
    
    br(),
    
    div(
      style = "background-color: #d4edda; padding: 15px; border-radius: 5px; border-left: 4px solid #28a745;",
      h4(style = "margin-top: 0; color: #155724;", "Scenario Interpretation"),
      p(strong("Severity Shock Impact: "), 
        ifelse(abs((sev_ibnr/base_ibnr - 1) * 100) < 5, "Low sensitivity",
               ifelse(abs((sev_ibnr/base_ibnr - 1) * 100) < 15, "Moderate sensitivity", "High sensitivity")),
        " - IBNR changes by ", round((sev_ibnr/base_ibnr - 1) * 100, 1), 
        "% when severity increases by ", input$severity_shock, "%"),
      p(strong("Development Speed Impact: "), 
        ifelse(abs((speed_ibnr/base_ibnr - 1) * 100) < 5, "Low sensitivity",
               ifelse(abs((speed_ibnr/base_ibnr - 1) * 100) < 15, "Moderate sensitivity", "High sensitivity")),
        " - IBNR changes by ", round((speed_ibnr/base_ibnr - 1) * 100, 1), 
        "% when development speeds up by ", input$dev_speed_shock, "%"),
      p(style = "margin-bottom: 0;", em("These scenarios help quantify the impact of adverse developments and support capital planning and risk assessment."))
    )
  )
})
```

# Chain Ladder Analysis
```{r chain-ladder-section}
conditionalPanel(
  condition = "output.show_sections",
  
  div(
    h2("Chain Ladder Analysis"),
    
    h3("Development Factors"),
    uiOutput("dev_factors_display"),
    
    h3("Ultimate Claims and Reserves"),
    uiOutput("reserves_display"),
    uiOutput("total_reserve_display")
  )
)
```
```{r run-model-reactive}
cl_results <- reactive({
  req(analysis_triggered())
  req(triangle_data())
  
  triangle <- triangle_data()
  cl_model <- MackChainLadder(triangle, est.sigma = "Mack")
  cl_summary <- summary(cl_model)
  
  list(model = cl_model, summary = cl_summary)
})
```
```{r dev-factors-display}
output$dev_factors_display <- renderUI({
  req(cl_results())
  
  cl_model <- cl_results()$model
  
  f_vec <- as.vector(cl_model$f)
  sigma_vec <- as.vector(cl_model$sigma)
  min_len <- min(length(f_vec), length(sigma_vec))
  
  dev_factors <- data.frame(
    Development_Period = paste(1:min_len, "to", 2:(min_len + 1)),
    Factor = f_vec[1:min_len],
    Std_Error = sigma_vec[1:min_len],
    Classification = ifelse(f_vec[1:min_len] > 1.20, "High",
                           ifelse(f_vec[1:min_len] >= 1.05, "Moderate", "Low/Mature"))
  )
  
  kable(dev_factors, 
        caption = "Age-to-Age Development Factors with Classifications",
        digits = 4,
        col.names = c("Development Period", "Factor", "Standard Error", "Classification"),
        align = c("l", "r", "r", "c")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE,
                  position = "center") %>%
    column_spec(4, bold = TRUE, 
                color = ifelse(dev_factors$Classification == "High", "#dc3545",
                              ifelse(dev_factors$Classification == "Moderate", "#ffc107", "#28a745"))) %>%
    HTML()
})
```
```{r reserves-display}
output$reserves_display <- renderUI({
  req(cl_results())
  
  cl_model <- cl_results()$model
  cl_summary <- cl_results()$summary
  
  reserves_df <- data.frame(
    Origin_Year = rownames(cl_summary$ByOrigin),
    Latest_Paid = cl_model$FullTriangle[, ncol(cl_model$FullTriangle)],
    Ultimate = cl_summary$ByOrigin$Ultimate,
    IBNR = cl_summary$ByOrigin$IBNR,
    Mack_SE = cl_summary$ByOrigin$Mack.S.E,
    CV = (cl_summary$ByOrigin$Mack.S.E / cl_summary$ByOrigin$IBNR) * 100
  )
  
  kable(reserves_df, 
        caption = "IBNR Reserves by Origin Year with Uncertainty Metrics",
        digits = c(0, 0, 0, 0, 0, 1),
        format.args = list(big.mark = ","),
        col.names = c("Origin Year", "Latest Paid", "Ultimate", 
                      "IBNR Reserve", "Std Error", "CV (%)"),
        align = c("l", "r", "r", "r", "r", "r")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE,
                  position = "center") %>%
    row_spec(nrow(reserves_df), bold = TRUE, background = "#f0f0f0") %>%
    HTML()
})
```
```{r total-reserve-display}
output$total_reserve_display <- renderUI({
  req(cl_results())
  
  cl_summary <- cl_results()$summary
  
  total_ibnr <- cl_summary$Totals["IBNR", ]
  total_se <- cl_summary$Totals["Mack S.E", ]
  cv <- total_se / total_ibnr * 100
  
  uncertainty_class <- ifelse(cv < 10, "LOW", ifelse(cv < 20, "MODERATE", "HIGH"))
  uncertainty_color <- ifelse(cv < 10, "#28a745", ifelse(cv < 20, "#ffc107", "#dc3545"))
  
  HTML(paste0(
    "<div style='background-color: #f8f9fa; padding: 20px; border-radius: 5px; margin-top: 20px;'>",
    "<h4>Total Reserve Summary</h4>",
    "<p><strong>Total IBNR Reserve:</strong> ", format(round(total_ibnr, 0), big.mark = ","), "</p>",
    "<p><strong>Standard Error:</strong> ", format(round(total_se, 0), big.mark = ","), "</p>",
    "<p><strong>Coefficient of Variation:</strong> ", round(cv, 2), "%</p>",
    "<p><strong>Uncertainty Classification:</strong> <span style='color: ", uncertainty_color, "; font-weight: bold;'>",
    uncertainty_class, "</span></p>",
    "<p style='font-size: 0.9em; color: #666; margin-top: 15px;'>",
    "Based on industry thresholds: Low (CV < 10%), Moderate (10% ≤ CV < 20%), High (CV ≥ 20%)",
    "</p>",
    "</div>"
  ))
})
```

# Visualizations
```{r visualizations-section}
conditionalPanel(
  condition = "output.show_sections",
  
  div(
    h2("Visualizations"),
    
    h3("Development Pattern"),
    plotOutput("development_plot", width = "100%", height = "700px"),
    
    h3("IBNR Reserves by Origin Year"),
    plotOutput("ibnr_plot", width = "100%", height = "600px"),
    
    h3("Paid Claims Heatmap"),
    plotOutput("heatmap_plot", width = "100%", height = "600px"),
    
    h3("Development Factor Trend"),
    plotOutput("factor_trend_plot", width = "100%", height = "600px")
  )
)
```
```{r development-plot-display}
output$development_plot <- renderPlot({
  req(cl_results())
  plot(cl_results()$model, lattice = TRUE)
})
```
```{r ibnr-plot-display}
output$ibnr_plot <- renderPlot({
  req(cl_results())
  
  cl_model <- cl_results()$model
  cl_summary <- cl_results()$summary
  
  reserves_df <- data.frame(
    Origin_Year = rownames(cl_summary$ByOrigin),
    IBNR = cl_summary$ByOrigin$IBNR,
    Mack_SE = cl_summary$ByOrigin$Mack.S.E
  )
  
  plot_df <- reserves_df[1:(nrow(reserves_df)-1), ]
  
  ggplot(plot_df, aes(x = Origin_Year, y = IBNR)) +
    geom_col(fill = "steelblue", alpha = 0.8) +
    geom_errorbar(aes(ymin = pmax(0, IBNR - Mack_SE), ymax = IBNR + Mack_SE),
                  width = 0.2, color = "red") +
    theme_minimal() +
    labs(title = "IBNR Reserves with Standard Error Bars",
         x = "Origin Year",
         y = "IBNR Reserve",
         subtitle = "Red bars indicate ±1 standard error range") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 10, color = "#666")) +
    scale_y_continuous(labels = scales::comma)
})
```
```{r heatmap-display}
output$heatmap_plot <- renderPlot({
  req(analysis_triggered())
  req(triangle_data())
  
  triangle <- triangle_data()
  triangle_matrix <- as.matrix(triangle)
  triangle_long <- data.frame(
    Origin = rep(rownames(triangle_matrix), times = ncol(triangle_matrix)),
    Development = rep(1:ncol(triangle_matrix), each = nrow(triangle_matrix)),
    Paid = as.vector(triangle_matrix)
  )
  
  ggplot(triangle_long, aes(x = Development, y = Origin, fill = Paid)) +
    geom_tile(color = "white", size = 0.5) +
    scale_fill_gradient(low = "lightblue", high = "darkblue", 
                        labels = scales::comma, na.value = "grey90") +
    theme_minimal() +
    labs(title = "Paid Claims Development Heatmap",
         x = "Development Period",
         y = "Origin Year",
         fill = "Paid Amount",
         subtitle = "Darker colors indicate higher claim amounts") +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
          plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 10, color = "#666"))
})
```
```{r factor-trend-display}
output$factor_trend_plot <- renderPlot({
  req(cl_results())
  
  cl_model <- cl_results()$model
  
  f_vec <- as.vector(cl_model$f)
  sigma_vec <- as.vector(cl_model$sigma)
  min_len <- min(length(f_vec), length(sigma_vec))
  
  dev_factors <- data.frame(
    Development_Period = paste(1:min_len, "to", 2:(min_len + 1)),
    Factor = f_vec[1:min_len],
    Std_Error = sigma_vec[1:min_len],
    Period = 1:min_len
  )
  
  ggplot(dev_factors, aes(x = Period, y = Factor)) +
    geom_line(color = "darkblue", size = 1.2) +
    geom_point(color = "darkblue", size = 3) +
    geom_ribbon(aes(ymin = Factor - Std_Error, ymax = Factor + Std_Error),
                alpha = 0.2, fill = "blue") +
    geom_hline(yintercept = 1.20, linetype = "dashed", color = "#dc3545", size = 0.8) +
    geom_hline(yintercept = 1.05, linetype = "dashed", color = "#ffc107", size = 0.8) +
    annotate("text", x = max(dev_factors$Period) * 0.9, y = 1.22, 
             label = "High Threshold (1.20)", color = "#dc3545", size = 3) +
    annotate("text", x = max(dev_factors$Period) * 0.9, y = 1.07, 
             label = "Maturity Threshold (1.05)", color = "#ffc107", size = 3) +
    theme_minimal() +
    labs(title = "Age-to-Age Development Factors with Standard Error",
         x = "Development Period",
         y = "Development Factor",
         subtitle = "Shaded area represents ±1 standard error; dashed lines show classification thresholds") +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 10, color = "#666")) +
    scale_x_continuous(breaks = 1:nrow(dev_factors),
                       labels = dev_factors$Development_Period) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
})
```

# Validation & Quality Checks
```{r validation-section}
conditionalPanel(
  condition = "output.show_sections",
  uiOutput("validation_display")
)
```
```{r display-validation}
output$validation_display <- renderUI({
  req(cl_results())
  req(triangle_data())
  
  triangle <- triangle_data()
  cl_summary <- cl_results()$summary
  
  total_ibnr_calc <- sum(cl_summary$ByOrigin$IBNR, na.rm = TRUE)
  total_ibnr_reported <- cl_summary$Totals["IBNR", ]
  ibnr_match <- abs(total_ibnr_calc - total_ibnr_reported) < 0.01
  
  cv_values <- (cl_summary$ByOrigin$Mack.S.E / cl_summary$ByOrigin$IBNR) * 100
  high_cv_count <- sum(cv_values > 20, na.rm = TRUE)
  
  dev_factors <- as.vector(cl_results()$model$f)
  outlier_dfs <- sum(dev_factors > 1.20, na.rm = TRUE)
  
  completeness_pct <- (sum(!is.na(triangle)) / (nrow(triangle) * ncol(triangle))) * 100
  
  div(
    h2("Validation & Quality Checks"),
    
    div(
      style = "background-color: #d4edda; padding: 20px; border-left: 4px solid #28a745; border-radius: 5px; margin-bottom: 20px;",
      h4(style = "margin-top: 0; color: #155724;", "✓ Automated Validation Results"),
      
      tags$table(
        class = "table table-sm",
        style = "width: 80%;",
        tags$thead(
          tags$tr(
            tags$th("Check", style = "text-align: left;"),
            tags$th("Result", style = "text-align: center;"),
            tags$th("Status", style = "text-align: center;")
          )
        ),
        tags$tbody(
          tags$tr(
            tags$td("IBNR Calculation Consistency"),
            tags$td(ifelse(ibnr_match, "PASS", "FAIL"), style = "text-align: center;"),
            tags$td(ifelse(ibnr_match, "✓", "✗"), 
                   style = paste0("text-align: center; color: ", 
                                ifelse(ibnr_match, "#28a745", "#dc3545"), "; font-weight: bold;"))
          ),
          tags$tr(
            tags$td("Data Completeness"),
            tags$td(paste0(round(completeness_pct, 1), "%"), style = "text-align: center;"),
            tags$td(ifelse(completeness_pct > 50, "✓", "⚠"), 
                   style = paste0("text-align: center; color: ", 
                                ifelse(completeness_pct > 50, "#28a745", "#ffc107"), "; font-weight: bold;"))
          ),
          tags$tr(
            tags$td("High Uncertainty Years (CV > 20%)"),
            tags$td(paste0(high_cv_count, " year(s)"), style = "text-align: center;"),
            tags$td(ifelse(high_cv_count == 0, "✓", "⚠"), 
                   style = paste0("text-align: center; color: ", 
                                ifelse(high_cv_count == 0, "#28a745", "#ffc107"), "; font-weight: bold;"))
          ),
          tags$tr(
            tags$td("Outlier Development Factors (DF > 1.20)"),
            tags$td(paste0(outlier_dfs, " period(s)"), style = "text-align: center;"),
            tags$td(ifelse(outlier_dfs == 0, "✓", "⚠"), 
                   style = paste0("text-align: center; color: ", 
                                ifelse(outlier_dfs == 0, "#28a745", "#ffc107"), "; font-weight: bold;"))
          )
        )
      ),
      
      if(high_cv_count > 0 || outlier_dfs > 0) {
        p(style = "color: #856404; margin-top: 15px; font-weight: bold;",
          "⚠ ATTENTION REQUIRED: ",
          ifelse(high_cv_count > 0, 
                 paste0(high_cv_count, " origin year(s) show high uncertainty (CV > 20%). "), ""),
          ifelse(outlier_dfs > 0,
                 paste0(outlier_dfs, " development factor(s) exceed threshold (>1.20). "), ""),
          "Review AI insights and Method Suitability sections for detailed analysis.")
      } else {
        p(style = "color: #155724; margin-top: 15px;",
          "All automated quality checks passed. Reserve estimates appear reasonable within expected parameters.")
      }
    ),
    
    h4("Recommended Next Steps"),
    tags$ul(
      tags$li(strong("Method Validation: "), "Consider comparing Chain Ladder results with Bornhuetter-Ferguson estimates as a reasonableness check"),
      tags$li(strong("Data Quality Review: "), "Verify source data accuracy and completeness for flagged origin years"),
      tags$li(strong("Expert Review: "), "Have senior actuaries review development factor outliers and high uncertainty years"),
      tags$li(strong("Assumption Testing: "), "Validate that portfolio characteristics (exposure, mix, claims handling) have remained stable"),
      tags$li(strong("Benchmarking: "), "Compare results against prior valuations, industry loss ratios, and actuarial expectations"),
      tags$li(strong("Documentation: "), "Archive this analysis with supporting documentation for regulatory compliance and audit trail"),
      if(input$run_scenarios) {
        tags$li(strong("Scenario Discussion: "), "Present scenario analysis results to stakeholders for capital planning and risk assessment")
      } else {
        tags$li(strong("Scenario Testing: "), "Enable scenario analysis (checkbox in Data Input section) to assess sensitivity to adverse developments")
      }
    )
  )
})
```

# Conclusion
```{r conclusion-section}
conditionalPanel(
  condition = "output.show_sections",
  uiOutput("conclusion_display")
)
```
```{r conclusion-display}
output$conclusion_display <- renderUI({
  req(cl_results())
  req(triangle_data())
  
  triangle <- triangle_data()
  cl_summary <- cl_results()$summary
  
  total_ibnr <- cl_summary$Totals["IBNR", ]
  total_se <- cl_summary$Totals["Mack S.E", ]
  cv <- total_se / total_ibnr * 100
  
  uncertainty_level <- ifelse(cv < 10, "low", ifelse(cv < 20, "moderate", "high"))
  uncertainty_color <- ifelse(cv < 10, "#28a745", ifelse(cv < 20, "#ffc107", "#dc3545"))
  
  HTML(paste0(
    "<div style='background-color: #f8f9fa; padding: 25px; border-radius: 5px;'>",
    
    "<p style='font-size: 1.1em;'>",
    "This comprehensive Chain Ladder reserving analysis demonstrates a rigorous, quantitative approach to IBNR estimation ",
    "with built-in validation, scenario testing, method suitability assessment, and AI-powered pattern recognition. ",
    "The total estimated IBNR reserve is <strong style='font-size: 1.2em;'>", 
    format(round(total_ibnr, 0), big.mark = ","), "</strong> ",
    "with a coefficient of variation of <strong>", round(cv, 2), "%</strong>, ",
    "classified as <strong style='color: ", uncertainty_color, ";'>", 
    toupper(uncertainty_level), " uncertainty</strong> based on industry-standard thresholds.",
    "</p>",
    
    "<h2>Key Findings</h2>",
    "<div style='background-color: white; padding: 15px; border-left: 4px solid #3498db; margin: 15px 0;'>",
    "<ul style='margin-bottom: 0;'>",
    "<li><strong>Scope:</strong> Analysis covers ", nrow(triangle), " origin years across ", 
    ncol(triangle), " development periods</li>",
    "<li><strong>Ultimate Projection:</strong> $", 
    format(round(cl_summary$Totals["Ultimate", ], 0), big.mark = ","), 
    " in total projected claims</li>",
    "<li><strong>Current Position:</strong> $", 
    format(round(cl_summary$Totals["Latest", ], 0), big.mark = ","), 
    " in paid claims to date</li>",
    "<li><strong>Reserve Ratio:</strong> ", 
    round(total_ibnr / cl_summary$Totals["Latest", ] * 100, 1), 
    "% of paid claims held in reserve</li>",
    if(input$run_scenarios) {
      paste0("<li><strong>Scenario Testing:</strong> Sensitivity analysis completed for severity (±", 
            input$severity_shock, "%) and development speed (±", input$dev_speed_shock, 
            "%) scenarios - see Scenario Analysis section</li>")
    } else {
      ""
    },
    "</ul>",
    "</div>",
    
    "<h2>Methodology & Professional Standards</h2>",
    "<p>This analysis employs the Mack Chain Ladder method with the following features:</p>",
    "<ul>",
    "<li><strong>Development Factors:</strong> Volume-weighted age-to-age progression with statistical classification</li>",
    "<li><strong>Uncertainty Quantification:</strong> Mack (1993) standard error estimation with threshold-based risk categorization</li>",
    "<li><strong>Method Suitability Assessment:</strong> Automated screening for red flags (immature data, volatile factors, exposure changes)</li>",
    "<li><strong>Pattern Recognition:</strong> AI-powered detection of outliers, trends, and concentration risks with actuarially correct interpretation</li>",
    "<li><strong>Validation Framework:</strong> Built-in consistency checks and data quality assessment</li>",
    "<li><strong>Scenario Analysis:</strong> Systematic stress testing for adverse development scenarios (illustrative, uniform shocks)</li>",
    "<li><strong>Alternative Methods:</strong> Documentation of when Bornhuetter-Ferguson or other methods should be considered</li>",
    "</ul>",
    
    "<h2>Important Limitations & Actuarial Judgment</h2>",
    "<div style='background-color: #fff3cd; padding: 15px; border-radius: 5px; margin: 15px 0;'>",
    "<p style='margin-bottom: 10px;'><strong>This analysis is subject to the following professional caveats:</strong></p>",
    "<ul style='margin-bottom: 8px;'>",
    "<li>Chain Ladder assumes stable development patterns and homogeneous portfolios — see Method Suitability section for assessment of these assumptions</li>",
    "<li>Development factors reflect <em>settlement speed and reporting patterns</em>, not claim magnitude — interpretation requires actuarial judgment</li>",
    "<li>Scenario analysis uses uniform, symmetric shocks — real adverse development may be asymmetric or concentrated in specific years</li>",
    "<li>Results should be compared with alternative methods (e.g. Bornhuetter-Ferguson) and validated against pricing expectations and industry benchmarks</li>",
    "<li>Final reserve selection requires professional actuarial judgment considering qualitative factors beyond this quantitative analysis</li>",
    "</ul>",
    "<p style='margin-bottom: 0; font-style: italic;'>",
    "<strong>Human Judgment Required:</strong> This analysis provides quantitative inputs to the reserving process. ",
    "Actuaries must exercise professional judgment in interpreting results, selecting final reserves, and communicating uncertainty to stakeholders.",
    "</p>",
    "</div>",
    
    "<h2>Audit Trail & Reproducibility</h2>",
    "<p style='background-color: #e7f3ff; padding: 15px; border-radius: 5px;'>",
    "All calculations are fully reproducible through the underlying R code using the ChainLadder package (version ", 
    as.character(packageVersion("ChainLadder")), "). ",
    "Development factors, ultimate projections, and uncertainty measures follow established actuarial standards (Mack 1993). ",
    "See the <strong>Assumptions & Methodology</strong> section for complete documentation of all parameters, ",
    "thresholds, and calculation methods. Source code available for audit upon request.",
    "</p>",
    
    "<hr style='margin: 25px 0;'>",
    
    "<p style='text-align: center; color: #666; font-style: italic;'>",
    "<strong>Report generated:</strong> ", format(Sys.time(), '%B %d, %Y at %I:%M %p'), "<br>",
    "<strong>Analyst:</strong> Rohith Narayanan<br>",
    "<strong>Method:</strong> Mack Chain Ladder with Method Suitability Assessment and AI-augmented Analysis<br>",
    "<strong>Framework:</strong> Quantitative inputs for actuarial judgment — final reserves require professional review",
    "</p>",
    
    "</div>"
  ))
})
```